\documentclass[12pt]{amsart}
\usepackage{fullpage}

\begin{document}

\begin{center}
Changes from version 1 to version 2
\end{center}

Example 2.2: The way the example written does in fact presume $K = F$. To cover the general case, one should appeal to Theorem 2.4.

Definition 2.5: Some remarks were added at the end to address the question about ``power series in $p$.''

Lemma 2.6: For essential surjectivity, ``tensoring`` with the splitting does not work, but there is a more sophisticated argument using the splitting that does give descent, based on a formal result of Joyal--Tierney.

Definition 2.8: The bijectivity for $\varphi$-modules comes from bijectivity on the underlying ring. We aren't sure whether any further comment is warranted.

Definition 2.17: We don't know of any significance to $Z_{L,A}$ being contained in an open affine subscheme besides the fact that it implies that $\widehat{Z}_{L,A}$ is affine.

Remark 2.19: We added some further discussion about the interplay between the schematic and adic Fargues--Fontaine curves, in order to illustrate the ``meaning'' of the spaces $P_{L,A,n}$ and how they relate to the extended Robba rings.

Remark 2.20, Lemma 2.21: These have been inserted to summarize the contact between perfectoid fields and older notions of ``deep ramification'' in the theory of local fields.

Definition 3.3: The valuation map is perhaps not most easily understood in terms of representations in tensors; we give a slightly different exposition. Similarly, while the degree map can indeed be constructed in terms of rational sections (this being the point of view taken in Fargues--Fontaine), it is easier in this context to express it in terms of $\varphi$-modules.

Lemma 3.4: We fleshed out slightly the ``induction'' comment.

Remark 4.9: This has been added to indicate the extent to which one can replace the cyclotomic tower, thus motivating the subsequent discussion.

Lemma 5.2: To alleviate confusion, we rewrote the statement to make it more clear what morphisms are giving rise to the quasi-isomorphisms in question.

Remark 5.8: We describe the analogue of Shapiro's lemma in more detail.

Remark 5.9: This remark has been added to clarify what happens in Theorem 5.7 when $K$ is not required to be finite over $\mathbb{Q}_p$.

Definition 6.1: Justification for the sheaf axiom (and acyclicity) added.

Definition 6.2: The group algebra is indeed formally of finite type over $\mathbb{Z}_p$. We added the standard reference for Berthelot's generic fiber construction (de Jong's paper on crystalline Dieudonn\'e theory), as well as some further explanation of the shape of $W_K$. The canonical representation is one-dimensional simply because it is acting on a one-dimensional projective module over $\mathcal{O}(W_K)$. Regarding the comment ``It would be nice to explain what this looks like in the case when $X$ itself is the generic rigid fiber of some formal scheme...'', it is not clear to us how to give such a description; it would depend on the module $M$, which is typically not easy to describe in terms of formal schemes.

Remark 7.4: To answer a referee's comment, we added some discussion of what $D_{\mathrm{crys}}$ means for a $(\varphi, \Gamma)$-module.


\end{document}